\chapter*{Abstract}

Low-power signal acquisition is a growing need in the Internet of Things and in distributed sensing, where many of the signals of interest are non-periodic. They appear as occasional, slowly changing events, such as a temperature reading that drifts over time, separated by long, quiet intervals. A conventional synchronous analog-to-digital converter is a poor match for this kind of input. Its global clock keeps the comparator bank and the clock distribution network switching at a fixed rate, so the converter dissipates dynamic power whether or not the input carries any new information.

This dissertation develops an asynchronous Flash ADC that spends power only when the input actually changes. The global clock is removed, and the conversion is driven by the signal itself, in a level-crossing manner, so the dynamic power follows the activity of the input and falls towards a static floor when the input is idle. Removing the clock also takes out the high-speed distribution network that, in a synchronous Flash, accounts for a large part of the power budget, and it shortens the conversion latency.

The converter is designed in a 16~nm CMOS process at a 0.8~V supply, with 4-bit resolution and a bank of fifteen comparison slicers. A complementary five-transistor operational transconductance amplifier, configured as a unity-gain voltage follower, reproduces the input across almost the whole supply range and drives the slicer bank. Each slicer is a current-mirror comparator whose decision level is fixed purely by the ratio of its transistor widths, so the fifteen thresholds are set by sizing rather than by a reference ladder. Those sizing-defined thresholds drift with process variation, so an offline body bias trimming strategy adjusts the bulk voltages of each slicer to bring the trip points back into place across the process corners, without adding any clocked activity or slowing the Flash conversion.

This work was carried out in an enterprise environment at Synopsys Portugal. It covers the theoretical analysis, the schematic-level design and the validation of the system through analog and mixed-signal co-simulation. The proposed converter is benchmarked against a synchronous Flash reference of the same resolution and technology, characterised under the same conditions. At a matched conversions-per-second workload, the asynchronous converter draws about a ninth of the power of the synchronous reference, and the saving grows when the input is quiet for most of the time, as the converter settles towards its idle floor.

\textbf{Keywords:} Analog-to-Digital Converter, Asynchronous Design, Flash ADC, Level-Crossing Sampling, Offset Trimming, Low Power, IoT.
