\chapter{Conclusions and Suggestions for Future Work}
\label{chap:Conclusions}

\section{Summary of the Work Developed}

This dissertation addressed the design and validation of an asynchronous Flash
Analog-to-Digital Converter intended for low-power, event-driven applications such as Internet
of Things sensor nodes and other sources of non-periodic signals. The central motivation for
this work stems from the fundamental inefficiency of conventional synchronous ADC
architectures when dealing with time-sparse signals: a global clock forces continuous switching
activity across the comparator bank and clock distribution network, dissipating dynamic power
regardless of whether the input signal is actually carrying new information. Removing that clock
and replacing it with a level-crossing, event-driven mechanism was the core hypothesis explored
throughout the dissertation.

The work began with a review of the state of the art in data conversion. The classical
ADC architectures (Flash, SAR, Pipeline and Sigma-Delta) were examined with particular
attention to the trade-offs between speed, resolution and power consumption. This analysis was
grounded in performance data from 140 publications drawn from ISSCC and JSSC, from which
Walden and Schreier figures of merit were extracted and compared across a wide range of sampling
frequencies and technology nodes. The review made clear that synchronous Flash converters
occupy the high-speed frontier but at the cost of a poor energy figure of merit, and that
asynchronous, event-driven operation is one of the most promising paths toward breaking this
constraint for sparse-signal applications. A dedicated discussion on the adaptation of standard
figures of merit to asynchronous systems was also included, since applying a fixed Nyquist
sampling frequency to a level-crossing converter is inherently inappropriate and leads to
misleading efficiency estimates.

The design of the asynchronous ADC itself evolved through a sequence of architectures.
The first candidate was a current-mirror flash configuration in which a transconductance cell
converts the differential input voltage into an error current, and a current-to-code converter maps
that current to a thermometer-coded output by exploiting asymmetric PMOS and NMOS mirror
ratios at each comparison node. This topology operates in continuous time and requires no
external clock, which made it a natural fit for the event-driven approach. The decision element of
this stage, the current comparator, was not assumed. Six published current comparator topologies
(current-mirror, Traff, Lu Chen, hysteresis, Min and Kim and CC-II) were redrawn at transistor
level and simulated in a common testbench, and the current-mirror cell was selected for having the
lowest propagation delay and the second-lowest average power of the six. The power and delay
figures reported for these comparators therefore come from direct simulation, not from estimates. The second architecture
investigated a complementary dual-input topology aimed at extending the input common-mode
range toward rail-to-rail operation by combining an NMOS and a PMOS differential pair with
cross-coupled output branches. Although conceptually attractive, this implementation could not be made
to work within the 0.8\,V supply budget of the target 16\,nm process. The mirror loads and the
cross-coupling stacked additional transistors in every signal path, the accumulated overdrive
requirements exceeded the available voltage headroom, and no feasible bias point could be found
that kept all devices in saturation simultaneously. A bulk-driven amplifier was then evaluated as a
low-voltage route to the same rail-to-rail input, but its positive-feedback transconductance boost
proved too sensitive to process corners and was discarded. The final architecture kept the
complementary rail-to-rail input but rebuilt it with two five-transistor OTAs, one with an NMOS
and one with a PMOS input pair, each loaded by a single active current mirror and biased through
the low-voltage tail mirror retained from the second architecture. With far fewer stacked devices, this
arrangement fits the available headroom, and the complementary OTA, configured as a unity-gain
voltage follower, drives a bank of CMOS slicer inverters through its low-impedance output. Each
slicer is sized with a distinct PMOS-to-NMOS transistor ratio, which sets its switching threshold
at a specific voltage level within the supply range. The ensemble of slicers produces an
asynchronous thermometer code whose transitions directly encode level-crossing events, with no
clock required at any point in the signal path.

Alongside the asynchronous core, a synchronous reference Flash ADC was implemented in the
same 16\,nm technology and under the same supply conditions to serve as a fair benchmark. This
reference uses fifteen StrongARM latch comparators driven by a ring oscillator with a tapered
buffer chain, and its power budget explicitly accounts for the clock distribution network, which is
precisely the overhead that the asynchronous design aims to eliminate. The comparison between the
two converters therefore isolates the effect of removing the global clock from an otherwise
structurally similar Flash architecture.

A critical practical challenge of the slicer-based thermometer coder is its sensitivity to process
variation. Since the threshold of each inverter is determined by the ratio of PMOS to NMOS drive
strength, any departure from the nominal device parameters shifts the trip points in a
bit-dependent manner, degrading the differential and integral non-linearity of the converter. To
address this, an offline body bias trimming strategy was designed and implemented using a
Verilog-A behavioural model. By applying distinct bulk voltages to the NMOS and PMOS
transistors of the slicer stage, the threshold voltage of each device type can be shifted through the
body effect, so the trip points can be steered back toward their nominal positions after
fabrication. Correction vectors were determined for the four standard process corners (SS, FF,
FS and SF) and were shown in simulation to improve the worst-case DNL at all four corners and
the worst-case INL at three of them, with the largest gains at the symmetric corners; the SF
corner proved the hardest to correct with a single global bias pair.

Finally, the complete system was validated through transistor-level analog and mixed-signal
co-simulation. The dynamic performance of the asynchronous converter was characterised using a
non-uniform to uniform resampling algorithm applied to the raw level-crossing output, enabling
standard spectral metrics such as SNDR and ENOB to be extracted and compared against both the
synchronous reference and representative state-of-the-art designs. The power measurements
confirmed the key theoretical prediction: dynamic power consumption scales linearly with input
signal frequency, and the converter enters a near-zero dynamic consumption state when the input
is idle. Measured at a matched conversions-per-second workload, the asynchronous core draws
$99.5$~\textmu W against the $886$~\textmu W of the complete synchronous reference, and the gap
widens for sparse inputs as the converter falls towards its idle floor. This power saving, rather
than a record Walden figure of merit, is where the energy advantage of the proposed converter
lies in the sparse-signal regime targeted by this work.

The contribution of this work lies less in any single block than in their combination:
ratioed-slicer quantisation, corner-indexed body bias trimming and clockless event generation,
brought together at $0.8$~V in a 16~nm process and benchmarked against a synchronous reference
of equal resolution and technology.

\section{Concluded Objectives}

The four specific objectives defined at the outset of this dissertation were all successfully
accomplished.

The first objective was the design of an asynchronous Flash ADC architecture that
dissipates dynamic power only when input signal events occur. This was achieved through the
final topology, a complementary five-transistor OTA driving the ratioed-slicer bank, which
operates in continuous time without any global clock signal. The theoretical analysis confirmed that the converter's dynamic power
scales linearly with the input signal frequency and approaches zero during idle intervals, directly
validating the intended activity-dependent power profile.

The second objective was the implementation of an offline trimming mechanism capable
of compensating the deterministic trip-point deviations of the comparison stage across process
corners. This was fulfilled through the body bias trimming strategy described in
Chapter~\ref{chap:trimming}, in
which pre-computed pairs of NMOS and PMOS bulk voltages are applied to the slicer bank to
restore the intended trip-point positions. The approach was validated across the SS, FF, FS, and
SF process corners, reducing the worst-case DNL at all corners and the worst-case INL at three
of the four, as reported in Table~\ref{tab:static_linearity_summary}.

The third objective was the validation of the complete system through analog and digital
co-simulation. The transistor-level netlist of the asynchronous ADC was simulated in a 16\,nm
CMOS process under a 0.8\,V supply, and both static linearity and dynamic performance metrics
were extracted. The FFT-based spectral analysis, performed on the uniformly resampled
level-crossing output, confirmed correct operation of the asynchronous core and the functionality
of the trimming logic, satisfying this objective.

The fourth objective was performance benchmarking against synchronous state-of-the-art
designs using established figures of merit. The proposed converter was compared against the
synchronous reference Flash ADC implemented under identical technology and supply conditions,
as well as against relevant designs from the ISSCC and JSSC literature, using the standard Walden
figure of merit with the rate fixed at twice the signal bandwidth. The proposed converter does not
set a record figure of merit, which is expected for a medium-resolution, megahertz-range Flash
converter when placed beside the high-resolution, kilohertz-bandwidth level-crossing designs
optimised purely for energy. Its advantage is instead in the matched-workload power, where it
consumes about a ninth of the power of the synchronous reference and scales down further for sparse
inputs, which is the operating regime targeted by this work.

Although process corner variation was successfully addressed by the
offline trimming strategy, Monte Carlo statistical variation was not corrected within the scope of
this work. The body bias trimming vectors were determined for the four deterministic process
corners and are therefore not adapted to the continuous distribution of device mismatch that Monte
Carlo analysis exposes. The Monte Carlo results of Chapter~\ref{chap:Benchmarking} quantify the
cost: the spread of a single code width reaches several LSB. This limitation is acknowledged as an important gap between the
presented prototype and a fully robust production-ready implementation, and it is the primary
motivation for the first suggestion for future work described in the following section.

\section{Future Work}

Two directions for future development are identified as the most impactful extensions of the
work presented in this dissertation.

\subsection{Analogue Process Corner Detection Approach}

The trimming strategy validated in this work relies on the correct identification of the active
process corner, which in the current implementation is supplied externally to the Verilog-A
behavioural model and used to select the appropriate pre-computed correction vector. In a
physical integrated circuit this external identification is not available: the fabricated die has no
direct means of knowing which combination of NMOS and PMOS device parameters it has been
assigned by the foundry.

A natural extension would therefore be a dedicated on-chip analogue process
detection circuit that senses the effective transistor characteristics of the fabricated device
and communicates the inferred process corner directly to the trimming logic. Such a circuit could
be implemented as a small reference cell: for example, a pair of matched inverter-based ring
oscillators, one biased toward NMOS-dominated operation and the other toward PMOS-dominated
operation, whose relative oscillation frequencies or propagation delays encode the current
process corner. The ratio between the two delay measurements maps unambiguously onto the
NMOS/PMOS balance of the process and distinguishes SS from FF and FS from SF corners. The
resulting analogue measurement would be digitised by a simple comparator and used to address a
lookup table of pre-computed body bias correction values, replacing the external corner
identification currently provided by the Verilog-A model. This approach would make the
trimming procedure fully self-contained and autonomous, removing the need for any external
process characterisation while preserving the static, low-overhead nature of the offline correction
strategy.

\subsection{Monte Carlo Trimming}

A more fundamental limitation of the current trimming approach is that it corrects deterministic
inter-corner shifts but leaves intra-die stochastic mismatch unaddressed. In practice, even at the
nominal process corner, individual transistors within the slicer bank will exhibit random threshold
voltage variations whose statistical distribution is characterised by Pelgrom's law \cite{pelgrom_1989}. These
variations are not captured by the four-corner trimming table, and their cumulative effect on DNL
and INL can be significant in a Flash converter where each comparator threshold must be
individually accurate.

Extending the calibration to handle Monte Carlo variation would require moving from a
corner-indexed lookup table to a converter-specific foreground calibration routine capable of
measuring the actual trip point of each slicer independently and computing an individual
correction for each one. One viable approach would be to inject a known ramp stimulus at the
converter input during a dedicated startup calibration phase and record the digital code transitions
to infer the actual trip point of each slicer. The deviation of each measured trip point from its
nominal target could then be mapped to the body bias correction required to bring it back into
specification, using the sensitivity relationship established in Chapter~\ref{chap:trimming}. Since the NMOS and
PMOS bulk voltages of each slicer would need to be individually controlled, this approach
requires the addition of a per-cell bias generator, which increases the area overhead relative to a
global corner correction. Nevertheless, the achievable linearity improvement and the
corresponding reduction in post-fabrication yield loss would likely justify this added complexity
in any application where the converter is produced at volume and must meet tight DNL
specifications across the full statistical distribution of process variability.