%% un-sdg.tex: UN Sustainable Development Goals (FEUP regulations)
%% -------------------------------------------------------

\chapter*{UN Sustainable Development Goals} \label{chap:UnitedNations}

The United Nations Sustainable Development Goals (SDGs) provide a global framework to achieve a better and more sustainable future for all. 
It includes 17 goals to address the world's most pressing challenges, including poverty, inequality, climate change, environmental degradation, peace, and justice. 

Electronic systems play a pivotal role in modern society, acting as the backbone for smart infrastructure, healthcare monitoring, and environmental sensing. However, the massive deployment of battery-operated devices poses a significant challenge regarding energy consumption and electronic waste. This dissertation contributes directly to the efficiency and sustainability of these systems.

The specific Sustainable Development Goals addressed by this work are:

\begin{description}
\item [SDG 7]
\textbf{Affordable and Clean Energy:} Ensure access to affordable, reliable, sustainable and modern energy for all. By optimizing the power consumption of data converters, we extend the battery life of devices and reduce the overall energy footprint of IoT networks.
\item [SDG 9]
\textbf{Industry, Innovation and Infrastructure:} Build resilient infrastructure, promote inclusive and sustainable industrialization and foster innovation. This work advances the state-of-the-art in microelectronics by proposing novel asynchronous architectures that enable smarter and more efficient industrial sensors.
\end{description}

\begin{center}
\begin{tabular}{|l|l|p{58mm}|p{52mm}|}
\hline
\textbf{SDG} & \textbf{Target} & \textbf{Contribution} & \textbf{Performance Indicators and Metrics} \\ 
\hline
\hline
    \multirow{3}{*}{7} & 7.3 & By double the global rate of improvement in energy efficiency, this work reduces the power waste in standby modes of electronic sensors. & Reduction in Energy per Conversion (fJ/conv) compared to synchronous architectures. \\
    \cline{2-4}
    & 7.a & Facilitate access to clean energy research and technology, including energy efficiency and advanced and cleaner fossil-fuel technology. & Development of ultra-low-power IP blocks for energy-harvesting systems. \\
\hline
    \multirow{2}{*}{9} & 9.5 & Enhance scientific research, upgrade the technological capabilities of industrial sectors, encouraging innovation. & Successful validation of the proposed calibration algorithm and circuit topology. \\
\hline
\end{tabular}
\end{center}